Testing Traversability of Protein Tunnels with Flexible Ligands Using Motion Planning

### 2017-06-28 paper.tex
corrections

\subsection{Computing initial configuration}

@@

\subsection{Computing single trajectory of the ligand}

@@

The trajectories are computed for all initial configurations $\qinit \in \QI$ and for each of the starting
configuration, multiple trajecotries are obtained.
For the purpose of this paper, we compute the tunnel characteristics using all computed trajectories.

A trajectory $P$ reaches the tunnel sphere $c_i \in t$ if the 3D Euclidean distances of the 
nearest configuration $q \in P$ towards $c_i$ is less than $r_i$ (radius of the sphere $c_i$).
Let $N(i)$ denote the number of trajectories, that reached $i-$th sphere of the tunnel, $i=1,\ldots,n$.
Three basic characteristics of the tunnel are computed from the trajectories: accessibility, throughput and the scaling factor.

The accessibility $A(i)$ of the sphere $i$ is the probability of reaching the sphere by the computed trajectories,
i.e., $A(i)=N(i) / N$, where $N$ is the total number of all trajectories.
The accessibility  shows how probable is to pass the tunnel upto the sphere $i$.
The  ligand passage may however be strongly affected by the first bottleneck, so the parts of the tunnel located behind
the bottleneck have low accessibility.

To better measure how probable is to pass a tunnel part, we should take into account only trajectories, that even have chance
to visit it, e.g. the trajectories that reached the preceding tunnel part.
The throughput $T(i)$ is the ratio of trajectories that passed sphere $i$ (i.e., visited sphere $i+1$) and reached the sphere $i$,
i.e, $T(i) = N(i+1) / N(i)$. 
The throughput is not computed for the last sphere ($i=N$).
The throughput shows a local accessibility of the tunnel parts.

@@

The proposed planner is allowed to scale down radii of the ligand up to the scale $\smin$. 
It can be expected, that narrow parts of the tunnels are more often passed with a more scaled down ligand, than
the wider parts.
Ligand scale $L(i)$ at the tunnel sphere $i$ is the average scale of ligand that reaches the sphere $i$.
The ligand scale is computed as the average value of scales of nearest configurations of the trajectories towards tunnel sphere
$i$.

@@

%In order to fursupport the tunnel exploration process performed by biochemists.
The computed properties of the tunnel are presented visually by mapping the accessibility, throughput and scale to the sphere representation
of the tunnel (Fig.~\ref{fig:properties}) using a color map.
Alternatively, the surface representation can be used to visualize the tunnel.
In this case, a 3D point on the tunnel surface is colored according to the property value in its nearest tunnel part $i$,  i.e., such part $i$ whose center $t_i$ is the closest to the point among all tunnel parts.

@@

The mapping using colors helps to identify difficult part of the tunnels, but it is also necessary to show the raw trajectories.
This is necessary for example if the trajectories detour from a part of the tunnel.
The na\"{i}ve visualization of all trajectories is possible, but too slow if the number of trajectories is in the order of thousands.

@@

To compare two trajectories (which are sequence of configuration in the configuration space), they are first converted
to a vector $v=(x_1,x_2,\ldots,x_M)$, where $x_i \in \R^3$ is a 3D
point representing the trajectory and $M$ is the parameter determining how precisely we want to represent the trajectories.
To obtain the vector $v$ for a given trajectory $P$, first a set of spheres 
 $S_1, ..., S_N$, where $S_i = (x_s, r_i)$ where $r_i = i \frac{|x_s x_e|}{N + 1}$.
%First, we represent each trajectory $t$ by a vector $v = (x_1, x_2, ..., x_N)$ of $N$ samples where $x_i \in \mathbb{R}^3$.
%In order to obtain the samples, we find the average starting point $x_s$ of all trajectories and the most distant (w.r.t. $x_s$) ending point $x_e$ also among all trajectories.
%Using $x_s$ and $x_e$ we define spheres $S_1, ..., S_N$ such that $S_i = (x_s, r_i)$ where $r_i = i \frac{|x_s x_e|}{N + 1}$.
Then, a vector $v$ for a trajectory $P$ is obtained such that an $i$-th sample $x_i$ in $v$ corresponds to the position in $P$ which is the nearest one to the surface of $S_i$.
Finally, we define a metric $d$ for trajectories $P_i$ and $P_j$ as $d(P_i, P_j) = \frac{1}{N} \sum_{1 \leq k \leq N} |x_k^i x_k^j|$.
The clustering is then computed using the UPGMA clustering algorithm~\cite{sokal1958statistical}.
When visualizing, we choose a representative from each cluster and depict its trajectory in 3D using a polyline.
Moreover, we map the representative's cluster size onto the width of the polyline.
In this manner, we are able to convey the information about all different trajectories together (see Fig.~\ref{fig:trajectories} right).

>>>>

\subsection{Computing Initial Configuration}

@@

\subsection{Computing Single Trajectory of the Ligand}

@@

The trajectories are computed for all initial configurations $\qinit \in \QI$ and for each of the starting
configuration, multiple trajectories are obtained.
For the purpose of this paper, we compute the tunnel characteristics using all computed trajectories.

Trajectory $P$ reaches the tunnel sphere $c_i \in t$, if the 3D Euclidean distance of the 
nearest configuration $q \in P$ towards $c_i$ is less than $r_i$ (radius of the sphere $c_i$).
Let $N(i)$ denote the number of trajectories that reached $i-$th sphere of the tunnel, $i=1,\ldots,n$.
Three basic characteristics of the tunnel are computed from the trajectories: accessibility, throughput, and the scaling factor.

The accessibility $A(i)$ of the sphere $i$ is the probability of reaching the sphere by the computed trajectories,
i.e., $A(i)=N(i) / N$, where $N$ is the total number of trajectories.
The accessibility  shows how probable is to pass the tunnel up to the sphere $i$.
The  ligand passage may however be strongly affected by the first bottleneck, so the parts of the tunnel located behind
the bottleneck have low accessibility.

To better measure the probability of passing a given tunnel part, we should take into account only trajectories that even have chance
to visit this part, e.g., the trajectories that reached the preceding tunnel part.
The throughput $T(i)$ is the ratio of trajectories that passed sphere $i$ (i.e., visited sphere $i+1$) and reached the sphere $i$,
i.e, $T(i) = N(i+1) / N(i)$. 
The throughput is not computed for the last sphere ($i=N$).
The throughput shows a local accessibility of the tunnel parts.

@@

The proposed planner is allowed to scale down the radii of ligand spheres up to the allowed scale $\smin$. 
It can be expected, that narrower parts of the tunnel are more often passed with a more scaled-down ligand than
the wider parts.
Ligand scale $L(i)$ at the tunnel sphere $i$ is the average scale of ligand that reaches the sphere $i$.
The ligand scale is computed as the average value of scales of the nearest configurations of the trajectories towards tunnel sphere
$i$.

@@

%In order to fursupport the tunnel exploration process performed by biochemists.
The computed properties of the tunnel are presented visually by mapping the accessibility, throughput, and scale to the sphere representation
of the tunnel (Fig.~\ref{fig:properties}) using a color map.
Alternatively, the surface representation can be used to visualize the tunnel.
In this case, a 3D point on the tunnel surface is colored according to the property value in its nearest tunnel part $i$,  i.e., such part $i$ whose center $t_i$ is the closest to the point among all tunnel parts.

@@

The mapping using colors helps to identify difficult parts of the tunnel, but it is also necessary to show the original trajectories.
This is necessary for example if the trajectories detour from a part of the tunnel.
The na\"{i}ve visualization of all trajectories is possible, but too slow if we are dealing with thousands of trajectories.

@@

To compare two trajectories (which are sequences of configurations in the configuration space), we first convert them
to a vector $v=(x_1,x_2,\ldots,x_M)$, where $x_i \in \R^3$ is a 3D
point representing the trajectory and $M$ is the parameter determining how precisely we want to represent the trajectories.
To obtain the vector $v$ for a given trajectory $P$, first a set of spheres 
 $S_1, ..., S_N$, where $S_i = (x_s, r_i)$ where $r_i = i \frac{|x_s x_e|}{N + 1}$.
%First, we represent each trajectory $t$ by a vector $v = (x_1, x_2, ..., x_N)$ of $N$ samples where $x_i \in \mathbb{R}^3$.
%In order to obtain the samples, we find the average starting point $x_s$ of all trajectories and the most distant (w.r.t. $x_s$) ending point $x_e$ also among all trajectories.
%Using $x_s$ and $x_e$ we define spheres $S_1, ..., S_N$ such that $S_i = (x_s, r_i)$ where $r_i = i \frac{|x_s x_e|}{N + 1}$.
Then, a vector $v$ for a trajectory $P$ is obtained such that an $i$-th sample $x_i$ in $v$ corresponds to the position in $P$ which is the nearest one to the surface of $S_i$.
Finally, we define a metric $d$ for trajectories $P_i$ and $P_j$ as $d(P_i, P_j) = \frac{1}{N} \sum_{1 \leq k \leq N} |x_k^i x_k^j|$.
The clustering is then computed using the UPGMA clustering algorithm~\cite{sokal1958statistical}.
When visualizing, we choose a representative from each cluster and depict its trajectory in 3D using a polyline.
Moreover, we map the representative's cluster size onto the width of the polyline.
In this manner, we are able to convey the information about all different trajectories together (see Fig.~\ref{fig:trajectories} right).
